---
title: "Task 1"
author: "Hannah Irish"
date: "2023-03-18"
output:
  html_document:
    code_folding: hide
---

# CA County Oil Spills 2008

### Include Packages
```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
library(tidyverse)
library(sf)
library(tmap)
library(here)
library(spatstat)
library(terra)

```


### Read in Data
```{r}

##read in spill data
oil_spills_sf <- read_sf(dsn=here("task1"), layer= "Oil_Spill_Incident_Tracking_[ds394]") %>%
  janitor::clean_names() %>%
  filter(inlandmari == "Inland") %>% ##filter by inland
  st_transform(crs = 32610) ##set CRS

##read in ca counties
ca_counties_sf <- read_sf(here("task1","data", "california_county_shape_file.shp")) %>%
  janitor::clean_names() %>%
  st_set_crs(4326) %>% ##set CRS 
  st_transform(32610)


```


### Set up data for tmpa
```{r}

spills_ppp <- as.ppp(oil_spills_sf) ###set ppp
ca_counties_win <- as.owin(ca_counties_sf) ###set window

spills_full <- ppp(spills_ppp$x, spills_ppp$y,
                  window= ca_counties_win) ###combine

```

### Rasterize

```{r}

spills_density <- density(spills_full, sigma=5000) ###set density

spills_r <- rast(spills_density) ##rasterize
crs(spills_r) <- crs(oil_spills_sf) ##make consistent CRS

###make tmap
tmap_mode("view") ##interactive
tm_shape(spills_r) +
  tm_raster(midpoint = NA,
            palette="Reds", ##colors
            legend.show = FALSE)
```


### Make choropleth
```{r}
###find number of spills in each county
count_oil_spills_sf <- oil_spills_sf %>%
  group_by(localecoun) %>%
  summarize(n_spills = n())

##rename county to name to set up for merge
colnames(count_oil_spills_sf)[1] = "name"

##make data frame to merge
spills_c <- as.data.frame(count_oil_spills_sf) %>%
  select(name, n_spills)


##merge
totaled <- merge(ca_counties_sf, spills_c, by.x="name", all.x=TRUE)

totaled[is.na(totaled)] <- 0


##Plot
ggplot() +
  geom_sf(data=totaled,
          size=0.5,
          aes(fill=n_spills))+ 
  scale_fill_continuous(low="lavenderblush1", high ="darkred")+
  labs(fill="Number of Spills in 2008", title="Oil Spills in 2008 by California County")+
  theme_void()
  

```

# Overview 
 
 The data has an overview of oil spills in CA in 2008, with timestamps and city and county information. There were the most oil spills in the year in the LA County, as you can see from the Choropleth map above. 
 
 Data souce:
 
 CA DFW. "Oil Spill Incident Tracking [ds394]" California.GOV. https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/explore?location=36.827189%2C-119.422009%2C6.74
 
 
 

